### 2019-10-30 Manuscript/Manuscript.tex
Updates from Overleaf

The values of four parameters within the silicate weathering component that modify the dependence of silicate weathering on temperature, runoff, and regolith thickness are poorly constrained. Rather than prescribing values for these parameters, we allow them and the Ca+Mg concentration of metamorphic and siliciclastic pixels to vary within reasonable bounds (see SI). This results in 93,600 unique combinations of these six parameters. For each combination, we compute spatially-resolved long-term \COtwo consumption associated with Ca+Mg fluxes using present-day runoff, temperature, and slope. We sum computed \COtwo consumption over watersheds for which data-constrained estimates are available \cite{Gaillardet1999a, Moquet2018a}, then calculate the coefficient of determination ($r^{2}$) between computed and measured \COtwo consumption in each of these watersheds. After eliminating parameter combinations that result in low $r^{2}$, 5,381 parameter combinations remain. The resulting global \COtwo consumption of these model runs all overlap with independently-derived estimates of global \COtwo outgassing flux as they should with the long-term carbon cycle being at steady-state (see SI).

Having calibrated the silicate weathering component of GEOCLIM, we use it to estimate the decrease in steady-state \pCOtwo associated with SEAI emergence. For the climate model component, we use temperature and runoff from a subset of the GFDL CM2.0 experiments \cite{Delworth2006b} (see SI). The primary strength of these experiments for this analysis is that all non-\COtwo forcings are held constant at values representative of pre-industrial conditions, allowing the effect of changing \pCOtwo on climatology to be isolated. To determine the position of SEAI paleoshorelines over the past 15~m.y., we use terrestrial and marine sedimentary deposits (Fig. \ref{fig:shoreline_growth}; see SI). The paleoshoreline data indicate that the Sunda-Banda Arc and New Guinea are primarily responsible for the increase in area since 15~Ma. Exhumation of the modern Sunda-Banda Arc is the result of ongoing arc-continent collision with the subducting Australian Plate \cite{Harris2006a}. Of the major islands in this arc, only Sumatra, Belitung, Bangka, Java, Bali, and Flores were subaerially exposed before 5~Ma. However, most of Sumatra and Java along with the non-volcanic islands of the Outer Banda Arc were elevated above sea level after 5~Ma, associated with slab roll-back and collision with the Australian continental margin \cite{Hall2013b}. In New Guinea, emergence in the Middle Miocene is associated with collision between the Melanesian Arc and Australia's distal margin \cite{vanUfford2005a}, which drove exhumation of the Irian-Marum-April Ophiolite Belt. Exhumation of the New Guinea Central Range accelerated over the past 4~m.y. associated with slab-breakoff and buoyant uplift while in eastern New Guinea, progressive jamming of the north-dipping subduction zone also caused major uplift over this time interval. These tectonic drivers and others throughout the region lead to progressive emergence over the past 15~m.y. that accelerated following 5~Ma (Fig. \ref{fig:shoreline_growth}B). This trend mirrors broad cooling over the Neogene that resulted in the initiation of Northern Hemisphere ice sheets (Fig. \ref{fig:shoreline_growth}C). We also include changes in areas of presently-submerged continental shelves such as the Sunda Shelf that were previously exposed (see SI).

We use GEOCLIM to estimate \pCOtwo associated with the reconstructed subaerial extent of SEAI at ca. 15, 10, and 5~Ma (``paleo-SEAI'' scenarios; Fig. \ref{fig:scenario_pCO2}). While we modify the land area, the position of the tectonic blocks remain fixed since we use a climate model forced with modern geography. While there has been motion of these tectonic blocks since 15~Ma, they have remained within tropical latitudes such that this fixed scenario is a good approximation of the paleogeography (see SI). We also test an end-member scenario, in which all islands associated with arc-continent collision in the region are removed (``removed SEAI'' scenario).

>>>>

The values of four parameters within the silicate weathering component that modify the dependence of silicate weathering on temperature, runoff, and regolith thickness are poorly constrained. Rather than prescribing values, we allow these parameters and the Ca+Mg concentration of metamorphic and siliciclastic pixels to vary within reasonable bounds (see SI). This results in 93,600 unique combinations of these six parameters. For each combination, we compute spatially-resolved long-term \COtwo consumption associated with Ca+Mg fluxes using present-day runoff, temperature, and slope. We sum computed \COtwo consumption over watersheds for which data-constrained estimates are available \cite{Gaillardet1999a, Moquet2018a}, then calculate the coefficient of determination ($r^{2}$) between computed and measured \COtwo consumption in each of these watersheds. After eliminating parameter combinations that result in low $r^{2}$, 5,381 parameter combinations remain. The resulting global \COtwo consumption of these model runs all overlap with independently-derived estimates of global \COtwo outgassing flux as they should with the long-term carbon cycle being at steady-state (see SI).

Having calibrated the silicate weathering component of GEOCLIM, we use it to estimate the decrease in steady-state \pCOtwo associated with SEAI emergence. For the climate model component, we use temperature and runoff from a subset of the GFDL CM2.0 experiments \cite{Delworth2006b} (see SI). The primary strength of these experiments for this analysis is that all non-\COtwo forcings are held constant at values representative of pre-industrial conditions, allowing the effect of changing \pCOtwo on climatology to be isolated.

To determine the position of SEAI paleoshorelines over the past 15~m.y., we use terrestrial and marine sedimentary deposits (Fig. \ref{fig:shoreline_growth}; see SI). The paleoshoreline data indicate that the Sunda-Banda Arc and New Guinea are primarily responsible for the increase in area since 15~Ma. Exhumation of the modern Sunda-Banda Arc is the result of ongoing arc-continent collision with the subducting Australian Plate \cite{Harris2006a}. Of the major islands in this arc, only Sumatra, Belitung, Bangka, Java, Bali, and Flores were subaerially exposed before 5~Ma. However, most of Sumatra and Java along with the non-volcanic islands of the Outer Banda Arc were elevated above sea level after 5~Ma \cite{Hall2013b}. In New Guinea, emergence in the Middle Miocene is associated with collision between the Melanesian Arc and Australia's distal margin \cite{Cloos2005a}, which drove exhumation of the Irian-Marum-April Ophiolite Belt. Exhumation accelerated over the past 4~m.y. in the New Guinea Central Range due to slab-breakoff and buoyant uplift, and in eastern New Guinea due to jamming of the north-dipping subduction zone \cite{Cloos2005a}. These tectonic drivers and others throughout the region lead to progressive emergence over the past 15~m.y. that accelerated following 5~Ma (Fig. \ref{fig:shoreline_growth}B). This trend mirrors broad cooling over the Neogene that resulted in the initiation of Northern Hemisphere ice sheets (Fig. \ref{fig:shoreline_growth}C). We also include changes in areas of presently-submerged continental shelves such as the Sunda Shelf that were previously exposed (see SI).

We use GEOCLIM to estimate \pCOtwo associated with the reconstructed subaerial extent of SEAI at ca. 15, 10, and 5~Ma (``paleo-SEAI'' scenarios; Fig. \ref{fig:scenario_pCO2}). Because we use a clmate model forced with modern geography, the position of the tectonic blocks remain fixed. Although there has been motion of these tectonic blocks since 15~Ma, they have remained within tropical latitudes such that this fixed scenario is a good approximation of the paleogeography (see SI). We also test an end-member scenario, in which all islands associated with arc-continent collision in the region are removed (``removed SEAI'' scenario).
